---
title: "Model without Western Blot"
output: html_document
---

Libraries/packages used:
```{r}
library(tidyverse)
library(varhandle)
library(R2jags)
library(coda)
library(xtable)
library(pROC)
library(aplot)
library(ROCit)
library(miscTools)
library(ggplot2)
library(plotROC)
```

Load csv file (created in "Echinococcus sero data.Rmd" file)
```{r}
dat <- read.csv("cleaned ovine sero-data.csv")
```

FULL MODEL - EXTENSION OF BASE MODEL TO INCLUDE NECROPSY DATA

Modify the data for the model - change column names, and result classification for western blot and necropsy data (0 = not infected, 1 = infected):
```{r}
model.dat3 <- dat %>% 
  select(X.1, diagnostico, X, westernblot) %>% 
  rename(n.sheep = X.1) %>% 
  rename(ELISA = X)

model.dat3$westernblot <- as.numeric(as.factor(model.dat3$westernblot))
model.dat3$westernblot[model.dat3$westernblot == 1] <- 0
model.dat3$westernblot[model.dat3$westernblot == 2] <- 1

model.dat3$diagnostico <- as.numeric(as.factor(model.dat3$diagnostico)) #hydatidosis is coded as number 4
model.dat3$diagnostico[model.dat3$diagnostico != 4] <- 0
model.dat3$diagnostico[model.dat3$diagnostico == 4] <- 1
```

Run the model:
```{r}
m.data3 <- list(N = 79,
                 ELISA = model.dat3$ELISA,
                diagnostico = model.dat3$diagnostico)

m.params3 <- c("P", "vsh", "vmn", "z", "status")

nc3 <- 2 #chains
ni3 <- 100000 #iterations
nb3 <- 10000 #burnin
nt3 <- 5 #thinning

set.seed(222)
result3 <- jags(model.file = "Model_without_WB.txt", data = m.data3, parameters.to.save = m.params3, n.iter = ni3, n.burnin = nb3, n.chains = nc3, progress.bar = "text", n.thin = nt3)
```

```{r}
result3.mcmc <- as.mcmc(result3, start = 10001, end = 100000, thin = 5)
save(result3, file = "model.Rdata")
load("model.Rdata")
```

Create a HTML document of diagnostic plots:
```{r}
mcmcplots::mcmcplot(result3.mcmc)
```

Produce a summary of the results:
```{r}
result3.sum <- (summary(result3.mcmc))
result3.sum
```

```{r}
stats_table2 <- as.data.frame(result3.sum$statistics)
quant_table2 <- as.data.frame(result3.sum$quantiles)

write.csv(stats_table2, "model_no_WB_stats.csv", row.names = T)
write.csv(quant_table2, "model_no_WB_BCI.csv", row.names = T)
```

Visualise the distribution of status mean values through density plots and histograms:
```{r}
status_mean2 <- as.vector(stats_table2$Mean[c(3, 14, 25, 36, 47, 58, 69, 80:81, 4:13, 15:24, 26:35, 37:46, 48:57, 59:68, 70:79)])

m3_plot <- ggplot(mapping = aes(status_mean2)) + geom_density() + labs(y = "Density", x = "Status value", title = "Model 2 status density") + theme(plot.title = element_text(hjust = 0.5, face = "bold"))
m3_plot

m3_hist <- ggplot(mapping = aes(status_mean2)) + geom_histogram(binwidth = 0.02, colour = "black", fill = "coral3") + labs(y = "Frequency", x = "Probability") + theme(plot.title = element_text(hjust = 0.5, face = "bold"))
m3_hist
```

```{r}
sp <- dat$diagnostico

sp <- ordered(sp, levels = c("hidatidosis", "cisticerco", "tysanosoma", "faciola", "dudoso", "negativo"))

status_mean_sp2 <- as.data.frame(cbind(as.numeric(status_mean2), sp))   

m3_hist_sp <- ggplot(status_mean_sp2, aes(x = status_mean2, fill = as.factor(sp))) + geom_histogram(binwidth = 0.02, position = "stack", alpha = 0.5, colour = "black") + labs(y = "Frequency", x = "Probability") + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), legend.text = element_text(size = 14), legend.title = element_text(size = 16), plot.margin=unit(c(1,1,1.5,1.2),"cm")) + scale_fill_discrete(name = "Necropsy Result", labels = c(expression(italic("E. granulosus")), expression(italic("Ta. hydatigena")),expression(italic("Th. actinioides")), expression(italic("F. hepatica")), "Indeterminate", "Negative"))
ggsave("status_other_nec.pdf", m3_hist_sp, dpi = 300, width = 300, height = 240, units = "mm")
m3_hist_sp
```