---
title: "Model without Western Blot"
output: html_document
---

Libraries/packages used:
```{r}
library(tidyverse)
library(varhandle)
library(R2jags)
library(coda)
library(xtable)
library(pROC)
library(aplot)
library(ROCit)
library(miscTools)
library(ggplot2)
library(plotROC)
```

Load csv file 
```{r}
dat <- read.csv("cleaned ovine sero-data.csv")
```

FULL MODEL - EXTENSION OF BASE MODEL TO INCLUDE NECROPSY DATA

Modify the data for the model - change column names, and result classification for western blot and necropsy data (0 = not infected, 1 = infected):
```{r}
dat$Necropsy <- as.numeric(as.factor(dat$Necropsy)) #Echinococcus is coded as number 4
dat$Necropsy[dat$Necropsy != 4] <- 0
dat$Necropsy[dat$Necropsy == 4] <- 1
```

Run the model:
```{r}
alt.data <- list(N = 79,
                 ELISA = dat$ELISA_OD,
                diagnostico = dat$Necropsy)

alt.params <- c("P", "vsh", "vmn", "z", "status")

nc2 <- 2 #chains
ni2 <- 100000 #iterations
nb2 <- 10000 #burnin
nt2 <- 5 #thinning

set.seed(222)
alt.result <- jags(model.file = "Model_without_WB.txt", data = alt.data, parameters.to.save = alt.params, n.iter = ni2, n.burnin = nb2, n.chains = nc2, progress.bar = "text", n.thin = nt2)
```

```{r}
alt.result.mcmc <- as.mcmc(alt.result, start = 10001, end = 100000, thin = 5)
save(alt.result, file = "alt_model.Rdata")
load("alt_model.Rdata")
```

Create a HTML document of diagnostic plots:
```{r}
mcmcplots::mcmcplot(alt.result.mcmc)
```

Produce a summary of the results:
```{r}
alt.result.sum <- (summary(alt.result.mcmc))
alt.result.sum
```

```{r}
stats_table2 <- as.data.frame(alt.result.sum$statistics)
quant_table2 <- as.data.frame(alt.result.sum$quantiles)

write.csv(stats_table2, "model_no_WB_stats.csv", row.names = T)
write.csv(quant_table2, "model_no_WB_BCI.csv", row.names = T)
```

Visualise the distribution of status mean values through density plots and histograms:
```{r}
status_mean2 <- as.vector(stats_table2$Mean[c(3, 14, 25, 36, 47, 58, 69, 80:81, 4:13, 15:24, 26:35, 37:46, 48:57, 59:68, 70:79)])

alt_plot <- ggplot(mapping = aes(status_mean2)) + geom_density() + labs(y = "Density", x = "Status value", title = "Alternative model status density") + theme(plot.title = element_text(hjust = 0.5, face = "bold"))
alt_plot

alt_hist <- ggplot(mapping = aes(status_mean2)) + geom_histogram(binwidth = 0.02, colour = "black", fill = "coral3") + labs(y = "Frequency", x = "Probability") + theme(plot.title = element_text(hjust = 0.5, face = "bold"))
alt_hist
```

```{r}
sp <- dat$Necropsy

sp <- ordered(sp, levels = c("Echinococcus", "Cysticercosis", "Thysanosoma", "Faciola", "Interderminate", "Negative"))

status_mean_sp2 <- as.data.frame(cbind(as.numeric(status_mean2), sp))   

alt_hist_sp <- ggplot(status_mean_sp2, aes(x = status_mean2, fill = as.factor(sp))) + geom_histogram(binwidth = 0.02, position = "stack", alpha = 0.5, colour = "black") + labs(y = "Frequency", x = "Probability") + theme(axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20), axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), legend.text = element_text(size = 14), legend.title = element_text(size = 16), plot.margin=unit(c(1,1,1.5,1.2),"cm")) + scale_fill_discrete(name = "Necropsy Result", labels = c(expression(italic("E. granulosus")), expression(italic("Ta. hydatigena")),expression(italic("Th. actinioides")), expression(italic("F. hepatica")), "Indeterminate", "Negative"))
ggsave("status_alt_nec.pdf", alt_hist_sp, dpi = 300, width = 300, height = 240, units = "mm")
alt_hist_sp
```